Windowed Least--Squares Reduced-Order Models for Dynamical Systems

### 2019-09-18 section3/general_form.tex
latest

\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
%\begin{equation}\label{eq:tclsrm}
%\begin{split}
%      &\underset{\statey \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\statey), \\
%      &\text{subject to } \;  \stateyArg{}{\timeStartArg{n}} =
%\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
%\end{split}
%\end{equation}

%\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$ as the solution to the minimization problem,
\begin{equation}\label{eq:tclsrm}
\begin{split}
       &\approxstateArgnt{n} =  \underset{\statey \in \stspace^n}{\text{arg\,min}} \; \mathcal{J}^n(\statey), \\ 
      &\text{subject to }\;  \approxstateArg{n}{\timeStartArg{n}} =
\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
\end{split}
\end{equation}
where $\mathbb{P}^n: \RR{N} \otimes \timeSpaceArg{n} \rightarrow \stspaceArg{n}$ is an orthogonal projection onto the 
trial space (e.g., $\elltwo$).
%\methodAcronym\ generates approximations to~\eqref{eq:FOM} by sequentially minimizing
%the FOM ODE residual over the trial space for each time slab.
%\methodAcronym\ is defined as follows,
%\begin{equation}\label{eq:tclsrm}
%\approxstate^n= \underset{\statey \in \stspace^n}{\text{arg\,min } }
%\mathcal{J}^n(\statey), \qquad n = 1,2,...,\nslabs,
%\end{equation}
% subject to
%the boundary condition, \begin{equation*} \approxstate^n(\timeStartArg{n}) =
%\begin{cases}\mathbb{P}^n \approxstate^{n-1}(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n \stateFOMIC & n=1, \end{cases} \end{equation*}
%where $\mathbb{P}^n: \RR{N} \rightarrow \stspaceArg{n}$ projects onto the 
%trial space (e.g., an $L^2$ projector).
\begin{remark}
It is possible to construct the trial spaces such that they have continuous support between time windows through the 
use of, e.g., affine transformations.
\end{remark}
%and 
%$\approxstate^n :  [\timeStartArg{n} , \timeEndArg{n}] \rightarrow
%\RR{\fomdim}$
%is the approximate solution over the $n$th time slab.% and $\approxstate_0 \in
%\RR{\fomdim}$ is the FOM initial condition
%projected onto the trial space; (e.g., via for $L^2$ projection).
We now define the trial spaces $\stspaceArg{n}$. We consider \spatialAcronym\ trial spaces and \spaceTimeAcronym\ trial spaces. 
We leave further investigation into other trial spaces, such as nonlinear manifolds~\cite{leeCarlberg}, as a subject for future work.
%A variety of trial spaces have been examined in the model-order reduction community.
%The most popular trial space comprises a linear, spatial-projection-only trial space (e.g., Eqns.~\eqref{eq:spatial_subspace}-\eqref{eq:spatial_subspace3}), although recently 
%nonlinear trial spaces~\cite{leeCarlberg} and space--time trial spaces~\cite{choi_stlspg,constantine_strom,2,3} have gained attention. This work focuses on the classic 
%spatial-projection-only trial space, and also briefly considers space--time trial spaces.
%We leave further investigation into other trial spaces as future work.

>>>>

\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
\begin{equation}\label{eq:tclsrm}
\begin{split}
      &\underset{\stateyArgnt{n} \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\stateyArgnt{n}), \\
      &\text{subject to } \;  \stateyArg{n}{\timeStartArg{n}} =
\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
\end{split}
\end{equation}
%\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$ as the solution to the minimization problem,
%\begin{equation}\label{eq:tclsrm}
%\begin{split}
%       &\approxstateArgnt{n} =  \underset{\statey \in \stspace^n}{\text{arg\,min}} \; \mathcal{J}^n(\statey), \\ 
%      &\text{subject to }\;  \approxstateArg{n}{\timeStartArg{n}} =
%\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
%\end{split}
%\end{equation}
where $\mathbb{P}^n: \RR{N} \otimes \timeSpaceArg{n} \rightarrow \stspaceArg{n}$ is an orthogonal projection onto the 
trial space (e.g., $\elltwo$).
%\methodAcronym\ generates approximations to~\eqref{eq:FOM} by sequentially minimizing
%the FOM ODE residual over the trial space for each time slab.
%\methodAcronym\ is defined as follows,
%\begin{equation}\label{eq:tclsrm}
%\approxstate^n= \underset{\statey \in \stspace^n}{\text{arg\,min } }
%\mathcal{J}^n(\statey), \qquad n = 1,2,...,\nslabs,
%\end{equation}
% subject to
%the boundary condition, \begin{equation*} \approxstate^n(\timeStartArg{n}) =
%\begin{cases}\mathbb{P}^n \approxstate^{n-1}(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n \stateFOMIC & n=1, \end{cases} \end{equation*}
%where $\mathbb{P}^n: \RR{N} \rightarrow \stspaceArg{n}$ projects onto the 
%trial space (e.g., an $L^2$ projector).
\begin{remark}
It is possible to construct the trial spaces such that they have continuous support between time windows through the 
use of, e.g., affine transformations.
\end{remark}
%and 
%$\approxstate^n :  [\timeStartArg{n} , \timeEndArg{n}] \rightarrow
%\RR{\fomdim}$
%is the approximate solution over the $n$th time slab.% and $\approxstate_0 \in
%\RR{\fomdim}$ is the FOM initial condition
%projected onto the trial space; (e.g., via for $L^2$ projection).
We now define the trial spaces $\stspaceArg{n}$. We consider \spatialAcronym\ trial spaces and \spaceTimeAcronym\ trial spaces. 
We leave further investigation into other trial spaces, such as nonlinear manifolds~\cite{leeCarlberg}, as a subject for future work.
%A variety of trial spaces have been examined in the model-order reduction community.
%The most popular trial space comprises a linear, spatial-projection-only trial space (e.g., Eqns.~\eqref{eq:spatial_subspace}-\eqref{eq:spatial_subspace3}), although recently 
%nonlinear trial spaces~\cite{leeCarlberg} and space--time trial spaces~\cite{choi_stlspg,constantine_strom,2,3} have gained attention. This work focuses on the classic 
%spatial-projection-only trial space, and also briefly considers space--time trial spaces.
%We leave further investigation into other trial spaces as future work.

### 2019-09-18 section3/tclsrm_formulation.tex
latest

\subsection{\methodAcronym\ Summary}
This section outlined the \methodAcronym\ method for model reduction. In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial space for a series of time windows:
\begin{equation*}
\approxstate^n= \underset{\statey \in \stspace^n}{\text{arg\,min } }
\mathcal{J}^n(\statey), \qquad n = 1,2,...,\nslabs,
\end{equation*}
where,
\begin{align*}
\mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
\RR{},
\end{align*}
subject to the boundary and continuity conditions. For the \spatialAcronym\ trial spaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
\begin{align*}\
&\mass \frac{d}{dt} \genstateArg{n}{t}  -  \basisspace^T
\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \mass \adjointArg{n}{t} , \\
 &\mass \frac{d}{dt} \adjointArg{n}{t}  + \basisspace^T \bigg[\frac{\partial
\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspace
 \adjointArg{n}{t} = \nonumber \\ 
&-\basisspace^T \big[
\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \bigg( \mathbf{I} -
\basisspace [\massArgt{n}{t}]^{-1} \basisspace^T \stweightingMatArgt{n}{t} \bigg)
 \bigg( \basisspace \frac{d}{dt}\genstateArg{n}{t}  -
\velocity(\veloargsromn) \bigg),  &
%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
\end{align*} 
subject to the boundary conditions~\eqref{eq:lspg_bcs1}-\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
trial spaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property.

>>>>

\subsection{\methodAcronym\ Summary}
This section outlined the \methodAcronym\ method for model reduction. In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial space for a series of time windows: \methodAcronym\  
computes the generalized coordinates $\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the minimization problems,
\begin{equation*}
\underset{\stateyArgnt{n} \in \stspace^n}{\text{minimize } }
\mathcal{J}^n(\stateyArgnt{n}), \qquad n = 1,2,...,\nslabs,
\end{equation*}
where,
\begin{align*}
\mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
\RR{},
\end{align*}
subject to the boundary and continuity conditions. For the \spatialAcronym\ trial spaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
\begin{align*}\
&\mass \frac{d}{dt} \genstateArg{n}{t}  -  \basisspace^T
\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \mass \adjointArg{n}{t} , \\
 &\mass \frac{d}{dt} \adjointArg{n}{t}  + \basisspace^T \bigg[\frac{\partial
\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspace
 \adjointArg{n}{t} = \nonumber \\ 
&-\basisspace^T \big[
\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \bigg( \mathbf{I} -
\basisspace [\massArgt{n}{t}]^{-1} \basisspace^T \stweightingMatArgt{n}{t} \bigg)
 \bigg( \basisspace \frac{d}{dt}\genstateArg{n}{t}  -
\velocity(\veloargsromn) \bigg),  &
%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
\end{align*} 
subject to the boundary conditions~\eqref{eq:lspg_bcs1}-\eqref{eq:lspg_bcs}. The Euler-Lagrange equations expose the fact that, for \spatialAcronym\
trial spaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property.
